---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![img](https://img.shields.io/badge/Lifecycle-Maturing-007EC6)](https://github.com/bcgov/repomountie/blob/master/doc/lifecycle-badges.md)
[![License](https://img.shields.io/badge/License-Apache%202.0-blue.svg)](https://opensource.org/licenses/Apache-2.0) 


```{r, echo=FALSE, results='asis', message=FALSE}
library("envreportutils")
```


# Analysis of trends and current amounts of protected lands and waters in B.C.

A set of R scripts to complete the analysis behind the Environmental Indicator [Protected Lands & Waters in B.C.](http://www.env.gov.bc.ca/soe/indicators/land/protected-lands-and-waters.html), published by [Environmental Reporting BC](http://www2.gov.bc.ca/gov/content?id=FF80E0B985F245CEA62808414D78C41B) in June 2016.

**Note**: Moving forward this analysis and indicator will be updated using the analysis and approach found [here](https://github.com/bcgov/land-designations-indicator).

## Usage

### Data

The analysis uses data from several sources. The data does not need to be pre-downloaded to run the code. 

-   [Canadian Protected and Conserved Areas Database         (CPCAD)](https://www.canada.ca/en/environment-climate-change/services/national-wildlife-areas/protected-conserved-areas-database.html)
-   [Ecoregions - Ecoregion Ecosystem Classification of British
    Columbia](https://catalogue.data.gov.bc.ca/dataset/d00389e0-66da-4895-bd56-39a0dd64aa78) (Licence: `r get_data_license("ogl-bc")`)
-   [Biogeoclimatic Ecosystem Classification (BEC)
    Map](https://catalogue.data.gov.bc.ca/dataset/f358a53b-ffde-4830-a325-a5a03ff672c3) (Licence: `r get_data_license("ogl-bc")`)
    
### Code

There are four core scripts that are required for the analysis; they need to be run in order:

- `01_load.R` - unzips the data files in the data folder
- `02_clean.R` - cleans and prepares the spatial data for analysis, especially resolving overlaps among types of protected area designations
- `03_analysis.R` - performs the spatial intersections and creates data summaries
- `04_plot.R` - creates plots for communicating the results

The file `fun.R` contains a few custom functions required for the analysis.

#### Required R packages:

Spatial packages:
```
sp
rgdal
rgeos
raster
maptools
geojsonio
bcmaps
rmapshaper
```

General data manipulation packages:
```
dplyr
tidyr
readr
```

Plotting packages:
```
ggplot2
RColorBrewer
ggthemes
envreportutils
```

Most packages used in the analysis can be installed from CRAN using `install.packages()`, but there are a few you will need to install using devtools:

```r
install.packages("devtools") # If you don't already have it installed

library(devtools)
install_github("bcgov/envreportutils")
install_github("bcgov/bcmaps")
```

#### Other required software

You will also need the [mapshaper](https://github.com/mbloch/mapshaper) node package, and for this you will require Node.js to be installed. 
You can install it from [here](https://nodejs.org/en/), then on the command line type:

```
npm install -g mapshaper
```


## Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an [issue](https://github.com/bcgov/bc_population_indicator/issues/).

## How to Contribute

If you would like to contribute, please see our [CONTRIBUTING](CONTRIBUTING.md) guidelines.

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

## License

    Copyright 2016 Province of British Columbia

    Licensed under the Apache License, Version 2.0 (the "License");
    you may not use this file except in compliance with the License.
    You may obtain a copy of the License at 

       http://www.apache.org/licenses/LICENSE-2.0

    Unless required by applicable law or agreed to in writing, software
    distributed under the License is distributed on an "AS IS" BASIS,
    WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
    See the License for the specific language governing permissions and
    limitations under the License.
    
This repository is maintained by [Environmental Reporting BC](http://www2.gov.bc.ca/gov/content?id=FF80E0B985F245CEA62808414D78C41B). Click [here](https://github.com/bcgov/EnvReportBC) for a complete list of our repositories on GitHub.
