---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![img](https://img.shields.io/badge/Lifecycle-Maturing-007EC6)](https://github.com/bcgov/repomountie/blob/master/doc/lifecycle-badges.md)
[![License](https://img.shields.io/badge/License-Apache%202.0-blue.svg)](https://opensource.org/licenses/Apache-2.0) 


```{r, echo=FALSE, results='asis', message=FALSE}
library("envreportutils")
```


# Analysis of trends and current amounts of protected lands and waters in B.C.

A set of R scripts to complete the analysis behind the Environmental Indicator [Protected Lands & Waters in B.C.](http://www.env.gov.bc.ca/soe/indicators/land/protected-lands-and-waters.html), published by [Environmental Reporting BC](http://www2.gov.bc.ca/gov/content?id=FF80E0B985F245CEA62808414D78C41B) in June 2016.

**Note**: Moving forward this analysis and indicator will be updated using the analysis and approach found [here](https://github.com/bcgov/land-designations-indicator).


### Data

The analysis uses data from several sources. The data does not need to be pre-downloaded to run the code. 

-   [Canadian Protected and Conserved Areas Database         (CPCAD)](https://www.canada.ca/en/environment-climate-change/services/national-wildlife-areas/protected-conserved-areas-database.html)
-   [Ecoregions - Ecoregion Ecosystem Classification of British
    Columbia](https://catalogue.data.gov.bc.ca/dataset/d00389e0-66da-4895-bd56-39a0dd64aa78) (Licence: `r get_data_license("ogl-bc")`)
-   [Biogeoclimatic Ecosystem Classification (BEC)
    Map](https://catalogue.data.gov.bc.ca/dataset/f358a53b-ffde-4830-a325-a5a03ff672c3) (Licence: `r get_data_license("ogl-bc")`)
    
### Code

Installation
This project leverages the `targets` package, a pipeline toolkit for data science projects in R. You can install `targets` from CRAN:

```{r}
#install.packages("targets")
```

Usage
Run `targets::tar_make()` to run project. This will run all of the analysis - no individual scripts are required. 

#### Required R packages:

The packages used in this analysis are catalogued in packages.R. It is likely that some of those packages won't be installed on your system. Those will need to be installed for this project to run.

#### Conserved Areas Shiny app

Part of the conserved areas indicator is a shiny app that details the percentage of conserved area by ecoregion. The shiny app can be opened after the targets workflow has been run via the following code: 
```{r}
#install.packages("shiny")
#shiny::runApp('app')
```


## Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an [issue](https://github.com/bcgov/bc_population_indicator/issues/).

## How to Contribute

If you would like to contribute, please see our [CONTRIBUTING](CONTRIBUTING.md) guidelines.

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

## License

    Copyright 2016 Province of British Columbia

    Licensed under the Apache License, Version 2.0 (the "License");
    you may not use this file except in compliance with the License.
    You may obtain a copy of the License at 

       http://www.apache.org/licenses/LICENSE-2.0

    Unless required by applicable law or agreed to in writing, software
    distributed under the License is distributed on an "AS IS" BASIS,
    WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
    See the License for the specific language governing permissions and
    limitations under the License.
    
This repository is maintained by [Environmental Reporting BC](http://www2.gov.bc.ca/gov/content?id=FF80E0B985F245CEA62808414D78C41B). Click [here](https://github.com/bcgov/EnvReportBC) for a complete list of our repositories on GitHub.

