---
topic: "land"
title: "Protected Lands & Waters in B.C."
output: envreportutils.internal::print_ver
---
<!--
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE}
library(ggplot2)
library(envreportutils)

root <- rprojroot::is_rstudio_project

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE)
load(root$find_file("tmp/out_gg.rda"))
load(root$find_file("tmp/out_data.rda"))
```


Protected lands and waters are legally established areas set aside to
conserve nature and cultural values. British Columbia's network of
protected lands and waters contribute to the global and Canadian targets
to set aside, by 2020, at least 17% of terrestrial areas and inland
water and 10% of coastal and marine areas through networks of protected
areas and other effective area-based conservation measures.^1,2^
This indicator reports on the trends in the establishment and status of
protected lands and waters in British Columbia.

-   **Protected lands and waters cover 15.4% of B.C.'s land base and
    3.2% of B.C.'s marine areas.** In the terrestrial realm, 15.0% is in provincial and federal parks and     protected areas, while the remaining 0.4% is in other provincial and federal designations such as Wildlife Management Areas and National Wildlife Areas, and private conservation lands.
-   The amount of recognized protected land in B.C. has more than
    doubled since the early 1990s.
-   The amount of protected land and water varies across
    British Columbia. For example, protected areas cover over 40% of the
    Gwaii Haanas, Chilcotin Ranges and Eastern Hazelton Mountains
    regions but less than 10% of the Fraser Basin, Fraser Plateau and
    Thompson-Okanagan Plateau regions.
-   Ecosystem representation varies within the protected areas system.
    For example, 30% of B.C.'s Boreal Altai Fescue Alpine habitats and
    5% of Coastal Douglas-fir habitats are protected.
-   Protected lands and waters, in particular [BC Parks](http://www.env.gov.bc.ca/bcparks/),
    make a major contribution to B.C.'s economy through job creation and
    park visitation &mdash; about 60% of British Columbians use a park every
    year.^3,4^
    
\newpage

### Map of Protected Lands & Waters in B.C. (2015)

Protected land and water designations include: all park and protected
areas managed by BC Parks, all national protected area designations,
such as National Parks and National Marine Conservation Areas,
provincially administered conservation lands such as Wildlife Management
Areas, and British Columbia's private conservation lands. Details on
specific designations and legal tools are provided below.

\vspace{16pt}

```{r prot-areas-map, fig.height=8, dev='png', dpi=300}
prot_map
```

\newpage

### Trends in the Protection of British Columbia's Terrestrial Ecoregions (1886-2015)

The map below shows how much land is protected within each of
B.C.'s 44 terrestrial ecoregions. The following series of plots show how that has changed over time.
[Ecoregions](http://www.env.gov.bc.ca/ecology/ecoregions/) are areas
with similar climate and geologic features.

\vspace{16pt}

```{r ecoregion-map, fig.height=8}
current_t_map + 
  labs(title = "Amount of Land and Water Protected in B.C. Ecoregions", 
       fill = "Percent of\nEcoregion Protected") +
  theme(plot.title = element_text(size = 12, face = "bold"))
```

```{r ecoregion-facet-line-plot, fig.height=12, fig.width=9}
ggplot(cum_summary_t_eco, 
       aes(x = prot_date, y = cum_percent_protected)) +
  geom_path(colour = "forestgreen") +
  facet_wrap(~ecoregion, labeller = label_wrap_gen(width = 20), ncol = 4) +
  scale_x_continuous(expand = c(0,0), breaks = function(x) round(seq(min(x),max(x), length.out = 5))) +
  scale_y_continuous(breaks = seq(0,100, length.out = 3)) +
  labs(x = "Year", y = "Cumulative Percent of Ecoregion Protected") +
  # theme_minimal() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 14, face = "bold"),
        panel.margin.x = unit(1, "lines"),
        plot.margin = unit(c(5,5,5,5),"mm"),
        axis.text.x = element_text(size = 9, angle = 45),
        strip.text = element_text(size = 12),
        strip.background = element_rect(fill = "white")) +
  geom_text(data = current_eco_t, x = 2001, y = 80,
            aes(label = paste0(round(cum_percent_protected, 1), "%")),
            size = 3) +
  ggtitle("Trends in Land and Water Protection in B.C. Ecoregions")
```


\newpage

**Interpreting the Trend Information:** The charts above do not reflect
changes to the size of *individual* protected areas over time. They
assume that each protected area was it's current (2015) size when it
was initally established. Only the terrestrial portion of the the
Georgia-Puget Basin ecoregion is represented on this map.

### Terrestrial Ecosystem Representation within Protected Lands in B.C.

-   Ecosystems in British Columbia are characterized as [biogeoclimatic
    zones](https://www.for.gov.bc.ca/hre/becweb/) &mdash; geographic areas with
    similar patterns of energy flow, vegetation and soils.
-   Protected lands are not evenly distributed among British
    Columbia's ecosystems. High elevation and mountainous areas such as
    the Spruce – Willow – Birch, Boreal Altai Fescue Alpine, and
    Interior Mountain-heather Alpine zones have more than the provincial
    average protected. Low elevation, warm, valley bottom areas tend to
    be underrepresented in the protected areas system &mdash; for example,
    Coastal Douglas-fir, Ponderosa Pine, and Interior Douglas-fir zones.
-   There is also high spatial variation in the proportion of ecosystem
    protection across the province. For example, most of the 20% Coastal
    Western Hemlock zone that is protected occurs along the west coast
    of Haida Gwaii. Protected areas in the Spruce – Willow – Birch zone
    are more evenly distributed.

\vspace{16pt}

```{r bec-bar-chart, fig.height=5}
zone_barplot +
  theme(plot.margin = unit(c(1,1,1,1), "lines"))
```

```{r bec-zone-map, fig.height=6}
bec_zone_map +
  theme(plot.title = element_text(size = 12, face = "bold"))
```

\newpage

### Marine Protected Areas in B.C.

-   Protected areas cover 3.2% of B.C.'s coastal and marine areas.
-   B.C. has 200 designated marine protected areas located in all six of
    the provincial marine ecoregions.
-   The [Endeavour Hydrothermal Vents Marine Protected
    Area](http://www.pac.dfo-mpo.gc.ca/oceans/protection/mpa-zpm/endeavour/index-eng.html)
    was established in B.C. waters in 2003 &mdash; the first marine protected
    area in Canada.

\vspace{16pt}

```{r marine-map-chart}
current_m_map + 
  theme(legend.position = "left", legend.direction = "vertical", legend.title = element_text(size = 13, face = "bold"), legend.text = element_text(size = 13), plot.margin = unit(c(1,1,1,1), "lines"))
```

\newpage

```{r marine-bar-chart}
summary_eco_m_plot +
  theme(plot.margin = unit(c(1,1,1,1), "lines"))
```

### Details of Protected Land & Water Designations in B.C.

-   **BC Parks (14.4% terrestrial; 0.9% marine)**
    -   Most of the protected land and a third of the protected waters
        in B.C. are managed by [BC Parks](http://www.env.gov.bc.ca/bcparks/).
    -   Protected areas managed by BC Parks are governed by several
        pieces of
        [legislation](http://www.env.gov.bc.ca/bcparks/aboutBCParks/legis.html)
        and fall into one of the following designations: Ecological
        Reserve, Provincial Park, Protected Area, Conservancy,
        Recreation Area.
-   **National Protected Areas (0.7% terrestrial; 2.2% marine)**
    -   The Government of Canada has established protected land and
        marine areas in British Columbia, for example National Parks,
        National Wildlife Areas, Marine Protected Areas and Migratory
        Bird Sanctuaries.
    -   These designations are governed by several pieces of federal
        legislation including: [*Canada National Parks Act*, *Canada
        National Marine Conservation Areas
        Act*](http://www.pc.gc.ca/eng/agen/dp-pd/lr-ar/lr-ar1.aspx),
        [*Canada Wildlife Act*, *Migratory Birds Convention
        Act*](http://www.ec.gc.ca/default.asp?lang=En&n=E826924C-1), and
        the [*Oceans
        Act (1996)*](http://laws-lois.justice.gc.ca/eng/acts/O-2.4/).
-   **B.C. Provincial Administered Conservation Lands (0.26%
    terrestrial; 0.07% marine)**
    -   [Administered Conservation
        Lands](http://www.env.gov.bc.ca/fw/habitat/conservation-lands/)
        are lands over which the Government of British Columbia has
        legal administrative and management authority.
    -   Wildlife Management Areas are Administered Conservation Lands
        that are designated under the provincial [*Wildlife
        Act*](http://www.env.gov.bc.ca/fw/wildlife/) to protect habitat
        for regionally, nationally or internationally
        significant species.
    -   Other types of Administered Conservation Lands are aquired by
        way of: [*Land
        Act*](http://www.bclaws.ca/EPLibraries/bclaws_new/document/ID/freeside/00_96245_01)
        section 106 Transfers of Administration, Fee Simple Acquisitions
        (through direct purchase, donation, or land exchange), and
        Lease Sites.
-   **B.C. Non-Government Organization Conservation Areas
    (0.1% terrestrial)**
    -   [B.C. Non-Government Organization Conservation
        Areas](http://ltabc.ca/resources/2012-02-05-22-20-02), also
        known as private conservation lands, include Fee Simple areas,
        defined as ownership of the title to a legal parcel, and
        Registerable Interests &mdash; other conservation securements registered
        on title, such as conservation covenants and leases.

**Note:** The sum of the numbers listed above will be greater than the
total provincial area reported on this page (15.4%). This is because
there is some overlap among designations. For example, there are about
10,000 hectares of land that are owned by non-governmental conservation
organizations, but are administered by the province. These lands are
included in both categories, but the overlap is removed in the overall
provincial and ecoregional analyses.

### Methods

This indicator uses data from 3 sources: the 2015 [Canadian Council on
Ecological Areas (CCEA) shapefile database of protected areas in
Canada](http://www.ccea.org/) accessed in January 2016; the provincially
administered conservation lands maintained by the [B.C. Ministry of
Forests, Lands and Natural Resource
Operations](http://www.env.gov.bc.ca/fw/habitat/conservation-lands/),
and data for privately held conservation lands, such as those owned by
land trusts, or private lands protected by conservation covenants,
available from the [British Columbia Non-Governmental Organization
Conservation Areas Database Working
Group](http://ltabc.ca/resources/2012-02-05-22-20-02). Municipal and
regional parks are not included in the analyses. The
[**R**](http://www.r-project.org) code for repeating the analyses
presented on this page is [available on
GitHub](https://github.com/bcgov/protected_areas_analysis). Overlaps
between protected areas have been accounted for in the analyses.

### References and Other Useful Links

[Canadian Environmental Sustainability Indicators: Canada's Protected
Areas](https://www.ec.gc.ca/indicateurs-indicators/default.asp?lang=en&n=478A1D3D-1)

[Canadian Council on Ecological Areas (CCEA)](http://www.ccea.org/)

[British Columbia Non-Governmental Organization Conservation Areas
Database Working Group](http://ltabc.ca/resources/2012-02-05-22-20-02)

[B.C. Government Conservation
Lands](http://www.env.gov.bc.ca/fw/habitat/conservation-lands/)

^1^[2020 Biodiversity Goals and Targets for
Canada](http://biodivcanada.ca/default.asp?lang=En&n=9B5793F6-1)

^2^[Convention on Biological Diversity: Aichi Biodiversity
Targets](https://www.cbd.int/sp/targets/)

^3^[BC Parks](http://www.env.gov.bc.ca/bcparks/) Facts and Figures:
[http://www.env.gov.bc.ca/bcparks/aboutBCParks/fact\_fig.html](/bcparks/aboutBCParks/fact_fig.html)

^4^Canada Parks Council. 2011. Economic Impact of Canada's National
Provincial and Territorial Parks in 2009. A Technical Report prepared by
The Outspan Group Inc. Amherst Island, Stella, Ontario. Available online
at: <http://www.parks-parcs.ca/english/cpc/economic.php>

### Data

\*By accessing these datasets, you agree to the license associated with
each file, as indicated in parentheses below.

-   [Indicator data: Summaries of Protected Lands and Waters
    in B.C.](https://catalogue.data.gov.bc.ca/dataset/6579744d-818c-4b26-839a-28a03edc0d98) (License: `r get_data_license("ogl-bc")`)
-   [Canadian Council on Ecological Areas (CCEA) shapefile database of
    protected areas in Canada](http://www.ccea.org/download-carts-data/)
-   [B.C. Government Conservation
    Lands](https://catalogue.data.gov.bc.ca/dataset/68327529-c0d5-4fcb-b84e-f8d98a7f8612) (License: `r get_data_license("ogl-bc")`)
-   [B.C. NGO Conservation Areas Database &mdash; Fee Simple
    Areas](http://ltabc.ca/resources/2012-02-05-22-20-02)
-   [Ecoregions - Ecoregion Ecosystem Classification of British
    Columbia](http://catalogue.data.gov.bc.ca/dataset/ecoregions-ecoregion-ecosystem-classification-of-british-columbia) (License: `r get_data_license("ogl-bc")`)
-   [Biogeoclimatic Ecosystem Classification (BEC)
    Map](http://catalogue.data.gov.bc.ca/dataset/biogeoclimatic-ecosystem-classification-bec-map) (License: `r get_data_license("ogl-bc")`)

Updated June 2016

Published and Available On-Line at Environmental Reporting BC (June 2016):
[](http://www.env.gov.bc.ca/soe/indicators/land/protected-lands-and-waters.html)

Email correspondence to: envreportbc@gov.bc.ca

\newpage

## Appendix: Breakdown of land designations in B.C.

**Note:** The sum of the numbers listed below will be greater than the
total provincial area reported on this page (15.4% for terrestrial, and 
3.2% for marine). This is because
there is some overlap among designations. For example, there are about
10,000 hectares of land that are owned by non-governmental conservation
organizations, but are administered by the province. These lands are
included in both categories, but the overlap is removed in the overall
provincial and ecoregional analyses.

### Terrestrial

\scriptsize

```{r}
bc_designations_summary$category <- gsub(" and ", " & ", 
                                         bc_designations_summary$category)

col_names <- c("Category", "Designation", "Total Area (Ha)", "Number", 
               "Percent of B.C.")

col_digits <- c(NA, NA , 2, 0, 4)

knitr::kable(bc_designations_summary[bc_designations_summary$BIOME == "T", 2:6], 
      col.names = col_names, row.names = FALSE, digits = col_digits)
```

\normalsize

### Marine

\scriptsize

```{r}
col_names <- c("Category", "Designation", "Total Area (Ha)", "Number", 
               "Percent of B.C.")

knitr::kable(bc_designations_summary[bc_designations_summary$BIOME == "M", 2:6], 
      col.names = col_names, row.names = FALSE, digits = col_digits)
```

\normalsize
